---
title: "Get started to GCubeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started to GCubeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(GCubeR)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.width = 7,
  fig.height = 5,
  message = FALSE,
  warning = FALSE
)

gembloux_kable <- function(x, caption = NULL, digits = 3) {
  # nombre de colonnes du data.frame (à garder avant le pipe)
  last_col <- ncol(x)

  pal_dark   <- "#004B87"  # dark blue
  pal_light  <- "#e8f5e9"  # light green
  pal_accent <- "#2e7d32"  # mid green

  x %>%
    kable(format = "html", digits = digits, caption = caption) %>%
    kable_styling(
      full_width        = FALSE,
      bootstrap_options = c("striped", "hover", "condensed")
    ) %>%
    row_spec(0, background = pal_dark, color = "white", bold = TRUE) %>%
    column_spec(1, background = pal_light, bold = TRUE) %>%
    column_spec(last_col, background = pal_accent, color = "white", bold = TRUE)
}
```


This vignette provides a quick introduction to the main functionalities of GCubeR using the example dataset included in the package, `GCubeR::data_rondeux`.

```{r echo=TRUE}
data("data_rondeux")
```


The dataset comes from a mixed, uneven-aged stand in the Bois Jacques Rondeux forest near Gembloux (Wallonia, Belgium). Tree diameters were measured at 1.50 m above ground and converted to circumference (`c150`). Total tree height was also measured, allowing the use of allometric equations that require both circumference and height (e.g., the Dagnelie two-entry equations).

```{r echo=FALSE}
data_rondeux %>% 
  ggplot(aes(x = c150, fill = species)) +
  geom_histogram() +
  theme_classic() +
  labs(x = "Circunference at 1m50 (in cm)",
       y = "Number of trees")
```

# Dagnelie equations

Most allometric equations require circumference measured at 1.30 m above ground (`c130`). Therefore, the first step is to convert `c150` into `c130`. The `c150_c130()` function automatically performs this conversion when supplied with a data frame containing the variables `c150` and the GCubeR species code.

```{r echo=TRUE}
data <- GCubeR::c150_c130(data_rondeux)
```

The Dagnelie one-entry equations estimate standing volume, \eqn{v_{c,22}} (m3 per tree), using only the circumference at 1.30 m (`c130`).

```{r}
data <- GCubeR::dagnelie_vc22_1(data)
```

Because the dataset also includes total tree height, we can use the Dagnelie two-entry equations, which estimate standing volume from both circumference at 1.30 m (`c130`) and total tree height (`htot`).

```{r echo=TRUE}
data <- GCubeR::dagnelie_vc22_2(data)
```


The predictions obtained from the one-entry and two-entry equations can then be compared.

For this dataset, the two-entry equations generally predict larger standing volumes than the one-entry equations, particularly for Quercus rubra. This difference likely reflects the fact that the inventories trees are taller than the average trees used to calibrate the one-entry equations, allowing the two-entry equations to better account for their observed height.

```{r echo=FALSE}
data %>% 
  ggplot(aes(y = dagnelie_vc22_2, x = dagnelie_vc22_1)) +
  facet_wrap(~species) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(slope = 1, intercept = 0,
              colour = "gray50", linetype = "dashed") +
  theme_classic() +
  labs(x = "Standing volume using 1-entry Dagnelie (dagnelie_vc22_1)",
       y = "Standing volume using 2-entry Dagnelie (dagnelie_vc22_2)")
```





