---
title: "Introduction to GcubeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to GcubeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GCubeR)
```
# Welcome to GcubeR

Forest quantification has become increasingly important in recent years.
Whether your goal is to anticipate financial flows, estimate standing timber value,
or quantify the carbon stored in your forest, working with tree volume equations
is rarely straightforward. Formulas are often difficult to find, coefficients vary
between species, and implementations can be tedious or error-prone.

GcubeR provides a unified and user-friendly toolbox that brings together the
most commonly used allometric equations in forestry. With only a few input
variables, the package handles all calculations for you.

GcubeR includes implementations of Algan, Rondeux, Dagnelie, and Vallet
volume equations, as well as CNIEFEB and Vallet's methods for estimating
biomass, carbon content, and CO₂ equivalent.

This vignette walks you through the workflow of applying GcubeR to your dataset
and shows how to extract the most useful information from your forest inventory.

---

# Global structure of GcubeR

## User input

To use **GcubeR**, the user must provide a data frame with at least:

- `c130`: stem circumference at 1.30 m height (cm)  
- `species_code`: full Latin species name in uppercase, e.g. `"QUERCUS_ROBUR"`

To be able to use **all** functions in the package, your data frame should contain:

- `c130`: stem circumference at 1.30 m height (cm)  
- `species_code`: species Latin name in uppercase (e.g. `"FAGUS_SYLVATICA"`)  
- `hdom`: dominant height (m)  
- `htot`: total tree height (m)  
- `dbh`: diameter at breast height (cm)

---

# Converting module

GcubeR provides tools to transform basic measurements so they fit the requirements
of the different volume and biomass equations:

- **`c150_c130`** : converts stem circumference measured at 1.50 m (`c150`)
  to circumference at 1.30 m (`c130`)
- **`add_c130_dbh`** : converts circumference at 1.30 m (`c130`) to
  diameter at breast height (`dbh`), and can also back-calculate `c130`
  from `dbh` when needed

---

# VTA module

The **VTA module** computes total aboveground volume (trunk + branches) using
three different methods.

### Bouvard method — `bouvard_vta`

Available for **QUERCUS** species.

Formula:

\[
v_{\mathrm{vta}} = 0.5 \times \left(\frac{dbh}{100}\right)^2 \times h_{tot}
\]

`dbh` is in cm, `htot` in m, result in m³ per tree.

---

### Algan method — `algan_vta_vc22`

Available for:

- `ABIES_ALBA`

Formula:

\[
v_{\mathrm{vta}} = 0.4 \times \left(\frac{dbh}{100}\right)^2 \times h_{tot}
\]

---

### Vallet method — `vallet_vta`

Available for the following species:

- `PICEA_ABIES`  
- `QUERCUS_ROBUR`  
- `FAGUS_SYLVATICA`  
- `PINUS_SYLVESTRIS`  
- `PINUS_PINASTER`  
- `ABIES_ALBA`  
- `PSEUDOTSUGA_MENZIESII`  

Formula:

\[
v_{\mathrm{vta}} =
\mathrm{form} \times \frac{1}{\pi \cdot 40000} \times c_{130}^2 \times h_{tot}
\]

where the *form factor* is species-specific and computed internally from
Vallet’s coefficients.

---
# VC22 module

The **VS22 module** computes the trunc volume until 22cm circumference using different
method

### non parametric équation

The non parametric equation are related on empirical volume table which gives them 
a better fitness for trees to evaluate trees from the same environement than the
one used for the model construction
---

### Dagnelie 1entry method — `dagnelie_vc22_1`

Available for the following species:
---


- `ABIES_ALBA`
- `ACER_CAMPESTRE`
- `ACER_PLATANOIDES`
- `ACER_PSEUDOPLATANUS`
- `AESCULUS_HIPPOCASTANUM`
- `ALNUS_GLUTINOSA`
- `ALNUS_INCANA`
- `BETULA_SP`
- `CARPINUS_SP`
- `CASTANEA_SATIVA`
- `CORLYUS_AVELLANA`
- `CRATAEGUS_SP`
- `CUPRESSUS_SP`
- `FAGUS_SYLVATICA`
- `FRAXINUS_EXCELSIOR`
- `JUNGLANS_SP`
- `LARIX_SP`
- `MALUS_SP`
- `PICEA_ABIES`
- `PICEA_SITCHENSIS`
- `PINUS_LARICIO`
- `PINUS_NIGRA`
- `PINUS_SYLVESTRIS`
- `POPULUS_TREMULA`
- `POPULUSxCANADENSIS`
- `PRUNUS_AVIUM`
- `PRUNUS_CERASUS`
- `PRUNUS_SP`
- `PSEUDOTSUGA_MENZIESII`
- `PYRUS_SP`
- `QUERCUS_PETRAEA`
- `QUERCUS_PUBESCENS`
- `QUERCUS_ROBUR`
- `QUERCUS_RUBRA`
- `QUERCUS_SP`
- `RHAMNUS_FRANGULA`
- `ROBINIA_PSEUDOACACIA`
- `SALIX_SP`
- `SAMBUCUS_SP`
- `SORBUS_ARIA`
- `SORBUS_AUCUPARIA`
- `TAXUS_BACCATA`
- `THUJA_PLICATA`
- `TILIA_SP`
- `ULMUS_SP`

# Function overview diagram
cat("
# Welcome to GcubeR

Forest quantification has become increasingly important in recent years.
Whether your goal is to anticipate financial flows, estimate standing timber value,
or quantify the carbon stored in your forest, working with tree volume equations
is rarely straightforward. Formulas are often difficult to find, coefficients vary
between species, and implementations can be tedious or error-prone.

**GcubeR** provides a unified and user-friendly toolbox that brings together
commonly used allometric equations in forestry. With only a few input variables,
the package handles all computations for you.

GcubeR includes implementations of Algan, Rondeux, Dagnelie, and Vallet
volume equations, as well as CNIEFEB and Vallet methods for estimating
biomass, carbon content, and CO\\(_2\\) equivalent.

This vignette walks you through the workflow of applying GcubeR to your dataset
and explains how to extract the most useful information from your forest inventory.

---

# Overall structure of GcubeR

## User input requirements

To use **GcubeR**, the user must provide a data frame containing at least:

- `c130`: stem circumference at 1.30 m height (cm)  
- `species_code`: full Latin species name in uppercase, e.g. \"QUERCUS_ROBUR\"

To access **all** functions in the package, the dataset should include:

- `c130` — circumference at 1.30 m  
- `species_code` — species Latin name  
- `hdom` — dominant height (m)  
- `htot` — total height (m)  
- `dbh` — diameter at breast height (cm)

---

# Conversion module

GcubeR provides tools to convert basic forest measurements as required by
different volume and biomass models:

- **`c150_c130`** — converts circumference measured at 1.50 m (c150)
  to circumference at 1.30 m (c130)
- **`add_c130_dbh`** — converts c130 to dbh, or dbh back to c130 when needed

---

# VTA module (Total Aboveground Volume)

The **VTA module** computes total aboveground volume (stem + branches)
using several classical forestry models.

### Bouvard method — `bouvard_vta`

Available for **QUERCUS** species.

The total aboveground volume is computed as:

\\[
v_{\\mathrm{VTA}} =
0.5 \\,\\times\\, \\left( \\frac{dbh}{100} \\right)^2 \\times h_{\\mathrm{tot}}
\\]

where:
- \(dbh\) is in centimeters,  
- \(h_{\\mathrm{tot}}\) is in meters.

---

### Algan method — `algan_vta_vc22`

Available for:

- `ABIES_ALBA`

The formula is:

\\[
v_{\\mathrm{VTA}} =
0.4 \\,\\times\\, \\left( \\frac{dbh}{100} \\right)^2 \\times h_{\\mathrm{tot}}
\\]

---

### Vallet method — `vallet_vta`

Available for the following species:

- `PICEA_ABIES`  
- `QUERCUS_ROBUR`  
- `FAGUS_SYLVATICA`  
- `PINUS_SYLVESTRIS`  
- `PINUS_PINASTER`  
- `ABIES_ALBA`  
- `PSEUDOTSUGA_MENZIESII`

Vallet's model computes total aboveground volume as:

\\[
v_{\\mathrm{VTA}} =
\\mathrm{form} \\,\\times\\,
\\frac{1}{\\pi \\cdot 40\\,000} \\,\\times\\,
c_{130}^2 \\,\\times\\, h_{\\mathrm{tot}}
\\]

where:
- \(c_{130}\) is circumference at 1.30 m (cm),  
- \(h_{\\mathrm{tot}}\) is total tree height (m),  
- \\(\\mathrm{form}\\) is a species-specific form factor estimated internally.

---

# VC22 module

The **VC22 module** computes stem volume down to a top circumference of 22 cm.

## Non-parametric equations

These equations rely on empirical volume tables, often providing better accuracy
for trees growing in conditions similar to those used to build the reference tables.

---

### Dagnelie 1-entry method — `dagnelie_vc22_1`

Available for the following species:

<div style=\"display:flex; gap:40px;\">

<div style=\"flex:1;\">

- `ABIES_ALBA`
- `ACER_CAMPESTRE`
- `ACER_PLATANOIDES`
- `ACER_PSEUDOPLATANUS`
- `AESCULUS_HIPPOCASTANUM`
- `ALNUS_GLUTINOSA`
- `ALNUS_INCANA`
- `BETULA_SP`
- `CARPINUS_SP`
- `CASTANEA_SATIVA`
- `CORLYUS_AVELLANA`
- `CRATAEGUS_SP`
- `CUPRESSUS_SP`
- `FAGUS_SYLVATICA`
- `FRAXINUS_EXCELSIOR`
- `JUNGLANS_SP`
- `LARIX_SP`
- `MALUS_SP`
- `PICEA_ABIES`
- `PICEA_SITCHENSIS`
- `PINUS_LARICIO`
- `PINUS_NIGRA`
- `PINUS_SYLVESTRIS`

</div>

<div style=\"flex:1;\">

- `POPULUS_TREMULA`
- `POPULUSxCANADENSIS`
- `PRUNUS_AVIUM`
- `PRUNUS_CERASUS`
- `PRUNUS_SP`
- `PSEUDOTSUGA_MENZIESII`
- `PYRUS_SP`
- `QUERCUS_PETRAEA`
- `QUERCUS_PUBESCENS`
- `QUERCUS_ROBUR`
- `QUERCUS_RUBRA`
- `QUERCUS_SP`
- `RHAMNUS_FRANGULA`
- `ROBINIA_PSEUDOACACIA`
- `SALIX_SP`
- `SAMBUCUS_SP`
- `SORBUS_ARIA`
- `SORBUS_AUCUPARIA`
- `TAXUS_BACCATA`
- `THUJA_PLICATA`
- `TILIA_SP`
- `ULMUS_SP`

</div>

</div>

---
")
